### 2018-12-10 submission/manuscript.Rmd
Working on Discussion

* Something like... our results suggest that the number of rounds and to a lesser degree the choice of DNA polymerase used in PCR impact the sequence data, the effects are consistent and are smaller than the biological differences between individuals

* be consistent within a study; do not interpret values like richness in absolute, but relative terms; take care in comparing results across studies

* GC, sequence length non-issues

* Other factors to investigate include impact of phylogenetic diversity, shearing, extension time, interactions of factors probably precludes ability to make definitive decisions.

* Interpret results in light of polymerase characteristics - Accuprime and Platinum have error rates that are ~10-times lower than that of Taq whereas Phusion, Q5, and Kappa are >100 times lower. These error rates track well with what we saw for sequencing error rates

* Minimize number of rounds of amplification - bias could be due to mis-determination of relative abundances by zymo

* Use multiple methods to triangulate on reality/validate claims

@@

Fecal samples showed a similar increase in median and range for the number of OTUs, based on number of cycles and polymerase used, as that found when analyzing mock community samples [Figure 4A]. However, there were few differences between the polymerases within each of the amplification cycles [Figure 4A]. Additionally, these differences did not translate to large bacterial community composition changes based on number of cycles or polymerase. The differences that were identified were small and occurred based on the number of cycles, independent of the polymerase used [Figure 4B]. There was also no OTU-based differences observed within the number of cycles between any of the polymerases for fecal samples. Interestingly, Random Forest models built on this OTU data had the highest classification success for individuals versus number of cycles or polymerase. In addition to the better model performance of Random Forest for individuals, the most important OTUs to this model were not the most important OTUs for the other models tested [Figure 5]. Collectively, these observations show that differences based on the number of cycles and between high fidelity DNA polymerases can be found. However, when using 16S rRNA gene sequencing these technical differences can be smaller than the biological differences between fecal bacterial communities within individuals.

The small differences observed due to error rate and chimera prevalence may be due to the actual DNA polymerase family being used within the PCR mix. DNA polymerases from distinct families have different binding affinities and error correction capacity [@polymerase_Ishino_2014]. However, based on the observations within our study this is unlikely to be the only contributor. Within our study the highest and lowest chimera prevalence both belonged to a family A polymerase (Platinum and Accuprime^TM^ respectively) [@polymerase_Ishino_2014]. Additionally, based on the information supplied by the respective manufacturers, the differences between the two PCR mixtures are not immediately apparent. Both PCR mixtures contain a recombinant *Taq* DNA polymerase, a *Pyrococcus* spp GB-D polymerase, and a platinum *Taq* antibody. Since it is not possible to know everything about the mixture beyond what was willingly provided by the manufacturer, it is possible that differences in how the recombinant *Taq* was generated or other compounds within the PCR mixture could be a contributing factor for the differences in error rate and chimera prevalence. Beyond the choice of the type of polymerase, there may be other ways to reduce the affect of polymerase dependent error rates and chimera prevalence on the downstream results.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Results have an impact on downstream diversity measurements due to effects on total OTUs

#Need to cover that different types of polymerase may affect outcome
#- Accuprime HiFi from recombinant A-Family Taq DNA polymerase, Pyrococcus spp GB-D polymerase (proof-reading), platinum Taq Antibody (inhibits room temp activity)

#- Kappa HiFi from B-family DNA polymerase (proof-reading baked in), propietary antibody (inhibit room temp activity)

#- Q5 HiFi from B-family thermostable DNA polymerase with 3´→ 5´ exonuclease activity, fused to a processivity-enhancing Sso7d domain to support robust DNA amplification

#-Phusion HiFi B-family DNA polymerase a novel Pyrococcus-like enzyme fused with a processivity-enhancing domain, generate blunt-ended products.

#- Platinum HiFi recombinant A-Family Taq DNA polymerase, Pyrococcus species GB-D polymeras, Platinum Taq Antibody

```

## Conclusion

>>>>

Our results suggest that the number of rounds of PCR and to a lesser degree the choice of DNA polymerase impact the analysis of 16S rRNA gene sequence data from bacterial communities. Although it was not possible to make direct connections between PCR conditions and specific sources of bias, we were able to identify general recommendations that reduce the amount of error, chimera formation, and bias. Researchers should strive to minimize the number of rounds of PCR and should use a high fidelity polymerase. Although specific PCR conditions impact the precise interpretation of the data, the effects were consistent and were smaller than the biological differences between individuals. Based on these observations, within a study, amplicons must be generated by consistent protocols to yield meaningful comparisons. When comparing across studies, values like richness, diversity, and relative abundances must be made in relative and not absolute terms.

The observed sequencing error rates and alpha diversity metrics followed the manufacturers' measurements of their polymerases' fidelity. Accuprime and Platinum have fidelity that are approximately 10-times higher than that of Taq whereas the fidelity of Phusion, Q5, and Kappa are more than 100 times higher. Among these polymerases, the Kappa polymerase resulted in the the lowest error rate, lowest chimera rate, and least bias across rounds of PCR. Considering polymerase development is an active area of commercial development and it likely that new polymerases will come to market, it is important for researchers to understand how changing the polymerase impacts downstream analyses for their type of samples.

### PCR bias
* GC, sequence length non-issues
* Other factors to investigate include impact of phylogenetic diversity, shearing, interactions of factors probably precludes ability to make definitive decisions.
* Minimize number of rounds of amplification - bias could be due to mis-determination of relative abundances by zymo

@@

### Conclusion
* Use multiple methods to triangulate on reality/validate claims
